Understanding the Nanoscale Structure of Inverted Hexagonal Phase Lyotropic Liquid Crystal Polymer Membranes

### 2018-10-08 Ben_Manuscripts/structure_paper/Final_Draft.tex
new comments
Responded to comments. Made axis labels bigger on plots.

  \begin{figure}[!htb]
        \centering
                \begin{subfigure}[t]{0.495\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated_words.pdf} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
	\begin{subfigure}[t]{0.405\linewidth}
                \centering
                \includegraphics[width=\textwidth]{SAXS.pdf}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
	\caption{(a) 2D-WAXS gives details about repeating features on the
		order of angstroms. Experimentalists have explained each of the 5 major
		reflections present as follows: (R-$\pi$) Aromatic head groups $\pi-\pi$ stack
		3.7 \AA~apart. (R-double) Monomers arrange vertically in a 2\textsubscript{1}
		helix. (R-alkanes) Alkane chain tails pack 4.5 \AA~apart. (R-spots) Monomer
		tails are tilted with respect to the membrane plane. (R-pores) As derived from
		SAXS, the pores are spaced 4.12 nm apart and pack hexagonally (b) (Reproduced
		from Ref.\citenum{feng_thin_2016}) The repeat spacing in the 1D-SAXS
		scattering pattern is characteristic of hexagonal packing. The leading peak,
		q*, represents the distance between the d\textsubscript{100} planes, which translates 
                into a distance between pore centers of 4.12 nm.} 
	\label{fig:SWAXS}
 \end{figure}

@@

@@

  We generated an ensemble of characteristic, low-energy vacuum monomer
  configurations by applying a simulated annealing process to a
  parameterized monomer. We cooled monomers from 1000 K to 50 K over 10
  ns. We randomly pulled a low energy configuration from the
  trajectory then reassigned charges using \texttt{molcharge}. Using the new
  charges, we annealed the monomer system again and pulled a random monomer
  configuration from the trajectory which we used for full system
  construction (Figure~\ref{fig:build_procedure}a). Section S-\ref{S-section:parameterization} 
  of the Supporting Information provides further detail of the parameterization process.

@@

  The time scale for self-assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  requiring a more efficient way to build the system. Previous work has shown
  a united-atom model of LLC can self-assemble into the H\textsubscript{II} phase
  configuration in $\sim$1000 ns \cite{mondal_self-assembly_2013}.  We
  attempted atomistic self-assembly by packing monomers into a box using Packmol
  \cite{martinez_packmol:_2009}. Simulations of greater than 100 ns show no
  indicators of progress towards an ordered system (see Section S-\ref{S-section:self_assembly} of the 
  Supporting Information). 
  To bypass the slow self-assembly process, we used a variety of automated procedures 
  to assemble monomers into a structure close to one of a number of hypothesized equilibrium configurations
  (Figure~\ref{fig:build_procedure}).
  %MRSF: where's the reference for the build scripts repository?
  %MRS13: Would be nice to say we put the scripts on GitHub.  Additional details about which scripts can be in supporting.
  These initial structures may result in equilibration into slowly-intercoverting
  metastable states, an issue we will address later in the paper.

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
  \includegraphics[width=\linewidth]{monomer_diagonal.pdf}
  \caption{}
  \end{subfigure}
  \begin{subfigure}[t]{0.42\textwidth}
  \includegraphics[width=\linewidth]{stacked_monomers.pdf}
  \caption{}
  \end{subfigure}
  \begin{subfigure}[t]{0.35\textwidth}
  \includegraphics[width=\linewidth]{initial_pore.pdf}
  \caption{}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
  \includegraphics[width=\linewidth]{initial_unit_cell.pdf}
  \caption{}
  \end{subfigure}
  %MRSF: I think this can be put in Supporting Information (we will get pushback to make this paper shorter, this is a good candidate)
  \caption{(a) We parameterized a single monomer and annealed it to produce a low-energy
		configuration. (b) We assembled monomers into columns by stacking them on top of each 
		other. (c) We duplicated each column and rotated them to create hydrophilic pore centers.
		We chose to stack twenty monomers into each column. (d) We duplicated the pores and 
		placed them into a monoclinic unit cell with hexagonal symmetry.}\label{fig:build_procedure}
  \end{figure}

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of columns of stacked monomers with periodic
  continuity along the pore axis, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. We prefer a small number of stacked
  monomers in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomers
  per column in order to minimize finite size effects with reasonable computational expense 
  and to obtain sufficient resolution when simulating XRD patterns (see further discussion
  in Section S-\ref{S-section:monomers_per_column} of the Supporting Information).

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. We find that the equilibrium
  configuration is sensitive to some while insensitive to others. For example, we find the starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure if one chooses
  a reasonable value. The initial distance between pores, within a wide range, also has 
  little effect on the equilibrated structure. However, one should not start them
  too close or there will be high energy repulsions during early equilibration. We chose a
  pore radius of 0.5 nm and an initial pore spacing of 4.5 nm, $\sim$10\% larger than the 
  experimental value of 4.12 nm for our initial configurations. A sensitivity analysis of 
  both parameters is presented in the Supporting Information, Section S-\ref{S-section:initial_config_dependence}. The 
  distance between vertically stacked monomers, the xy position of monomers with respect 
  to vertically adjacent monomers, and the number of columns per pore do influence the 
  equilibrated structure and require further justification for their choices. We rely on 
  experimental data to inform them.

  We chose the vertical spacing between monomers for the initial configuration based
  on R-$\pi$ and then allowed the system to readjust during equilibration. We rotated 
  each monomer so the plane of its aromatic head group would be coplanar with the xy plane. We
  explored three different initial monomer spacings. The first is exactly
  equal to R-$\pi$ with monomers placed so aromatic rings stack 3.7 \AA~apart in
  the $z$-direction. We also extensively explored a second system with an initial spacing of 5
  \AA. We briefly explored a third system with an initial spacing of 10
  \AA. However this largest spacing yields non-physical behavior which is detailed in the 
  Supporting Information, Section S-\ref{S-section:initial_dbwl}.

  We chose the relative orientation between vertically adjacent monomers in each column 
  based on clues from diffraction data as well as the various known stacking modes of 
  benzene and substituted benzene rings: sandwiched, parallel-displaced and T-shaped
  \cite{sinnokrot_estimates_2002}. We ruled out the T-shaped configuration
  because its $\sim$5 \AA~equilibrium stacking distance \cite{sinnokrot_estimates_2002}
  is inconsistent with R-$\pi$. It is also infeasible for the monomers to orient in the 
  T-shaped conformation because of the bulky tail groups. We explored the system's 
  preference towards the sandwiched vs. parallel displaced stacking modes in some detail.
  Both have reported stacking distances near the R-$\pi$ value of 3.7 \AA. Head groups in
  our sandwiched initial configuration stack directly on top of each other while
  head groups in the parallel displaced initial configuration stack with an offset
  of $180^\circ/ncol$ where $ncol$ is the number of columns per pore. See Figure 
  S-\ref{S-fig:stacking} of the Supporting Information for a detailed illustration 
  of the initial configurations in each mode.

@@

  We developed equilibration schemes for creating dry (i.e. thermotropic LC phase)
  and wet (i.e. LLC phase) configurations.
  Both schemes start with an initial configuration generated according to the
  previous guidelines. For wet systems, we added the desired concentration of
  water to the initial configuration and carried out equilibration in the same
  way as the dry systems. First, we fixed monomer head groups in place using
  position restraints with a force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We
  gradually released the position restraints by decreasing the force constants
  over a series of NVT simulations. We allowed the resulting unrestrained
  structure to equilibrate for 5 ns in the NPT ensemble with pressure controlled
  by the Berendsen barostat, followed by NPT equilibration simulations run for at
  least 400 ns using the Parrinello-Rahman barostat. More equilibration details
  are given in Section S-\ref{S-section:equilibration} of the Supporting Information.

@@

  To calculate the equilibrated pore spacing, we measured the distance between
  pore centers. We located the pore centers by averaging the coordinates of sodium
  ions in their respective pores. We generated pore spacing statistics 
  using the bootstrapping technique (See Section S-\ref{S-section:p2p_stats} of the
  Supporting Information). 
  The pore spacings calculated in this way are consistent
  with one half of the $x$ and $y$ box vectors (See, Table S-\ref{S-table:p2p} in the
  Supporting Information). However, our method for their calculation does a better 
  job capturing the spread of pore-to-pore distances.

@@

  We measured the intensity of R-pores based on the intensity of the
  d\textsubscript{100} peak (the leading peak closest to $q_r$ = 0) of the $q_r$
  cross-section of the patterns at $q_z$=0 (Figure~\ref{fig:saxs_xsection}). The
  beamstop covers most of the small angle reflections in the experimental 2D-WAXS
  pattern. In order to compare the simulated intensity of R-pores to experiment,
  we used the $q_r$ cross-section of 2D-SAXS which was generated from the same
  sample (Figure S-\ref{S-fig:2DSAXS} of the Supporting Information). Since the d\textsubscript{200} peak is
  partially exposed in the experimental 2D-WAXS pattern, we normalized the
  experimental 2D-SAXS cross-section by matching the intensity of the
  d\textsubscript{200} peak between it and the experimental 2D-WAXS
  cross-section. It is possible that the peak is not fully captured in the 2D-WAXS
  pattern and that we have underestimated the intensity of R-pores in the
  experimental pattern.

@@

  We explored the pores' compositions by measuring the average number densities
  of various monomer components as a function of distance from the pore centers.
  We looked at the average number density of sodium ions, aromatic rings and 
  carbon atoms making up the monomer tails. We binned the radial distance of all
  atoms in each group from the pore centers, then normalized by the volume of the
  annulus defined by the bin edges and the z box vector (See Figure~S-\ref{S-fig:rdf_diagram}
  of the Supporting Information).

@@

  We calculated ionic conductivity using the Nernst-Einstein relationship, which 
  relates the DC ionic conductivity, $\sigma$, to ion diffusivity, $D$, 
  concentration, $C$, ion charge, $q$, the Boltzmann constant, $k_b$, and 
  absolute temperature, $T$: 
  \begin{equation}
	\sigma = \dfrac{q^2CD}{k_b T} 
	\label{eqn:nernst_einstein}
  \end{equation}
  We measured sodium ion diffusion coefficients by calculating the slope
  of the linear region of the $z$-direction mean square displacement curve as
  indicated by the Einstein relation \cite{einstein_investigations_1956}. We
  visualized the MSD plot to determine where to begin and end a linear fit. We
  measured ion concentration with respect to the volume of the entire unit cell. 
  More details are provided in the Supporting Information, Section~S-\ref{S-section:ionic_conductivity}.

@@

  We based our cross-linking algorithm on the known reaction mechanism 
  (Figure S-\ref{S-fig:xlink_mech} of the Supporting Information). The reaction 
  takes place at the terminal vinyl groups
  on each alkane tail. The procedure is carried out iteratively. Each iteration, the
  algorithm chooses carbons to cross-link based on the distance between eligible 
  carbon pairs. The algorithm then updates the topology with the new bonds and atom
  types, energy minimizes the system and runs a short simulation before selecting 
  the next group of eligible carbons atoms.

  \subsection{Source code}
  %MRSF: Suggest moving these references to where each code is described instead of separately here.
  %BJC: the hyperref package really messes up the labels
  The python scripts used to simulated XRD patterns are publicly available online at \\
  \texttt{https://github.com/joeyelk/MD-Structure-Factor}. Python scripts used to conduct
  all other post-simulation trajectory analysis, are also available online at \\ 
  \texttt{https://github.com/bencoscia/llcsim}.
  %BJC: llcsim repository only contains necessary scripts. Should probably move to Shirts-group repo

@@

  Our simulations best support a model built with 5 monomer columns per pore based on
  the measured equilibrated pore-to-pore distances. To identify the density of
  monomer columns around each pore, addressing question \ref{point:monomernum} in the introduction, we ran
  simulations of systems created with 4--8 columns per pore. We built systems in both
  the parallel displaced and sandwiched configurations and equilibrated them according
  to the dry equilibration procedure. We tested all systems with an initial vertical monomer 
  spacing, $\mathit{d}$, of 3.7 \AA~in accordance with R-$\pi$. We tested 4 additional 
  systems with monomers initially spaced 5 \AA~apart vertically (see Section~\ref{S-section:initial_pore_spacing}
  of the Supporting Information for more details on sensitivity to initial monomer spacing). 
  We considered the pore-to-pore spacing to be equilibrated as defined in Section~\ref{method:equil_time}.
  Figure~\ref{fig:p2p} shows the equilibrated pore-to-pore distances for all systems tested.

@@

  All systems tested, although equilibrated from the perspective of the metrics used here, are
  frozen in metastable basins. Not all make physical sense or fit the experimental profile that 
  we are trying to match. In the limit of infinite simulation time, all systems will in theory 
  converge to a single equilibrium configuration, but that time is far beyond the 100's of 
  nanosecond simulated here. For simplicity, we group the systems studied here into the ordered
  and disordered basins. What we find is that any system equilibrated with $\mathit{d}=3.7$ \AA~can generally 
  be characterized as being in a more ordered basin, and any system equilibrated with $\mathit{d}=5.0$ \AA~can
  be characterized as begin in a more disordered basin. The extra space between stacked monomers
  in the disordered basin systems gives the monomer head groups more rotational freedom. We 
  quantified the ordering of the head groups using the nematic order parameter (see Section 
  S-\ref{S-method:nematic_order} for details of the calculation). Disordered basin systems have
  a lower nematic order parameter (meaning they are more disordered) than ordered basin systems. 
  Generally, when monomers are started further apart, they stay further apart than systems where
  monomers are started closer together (see Table~\ref{table:correlation_length}). This increased
  equilibrated stacking distance of monomers in disordered basin systems increases their discrepancy
  with experimental stacking distances derived from WAXS relative to ordered basin systems 
  (See section~\ref{section:rpi}).

@@

  We can better understand the structure of the system by simulating XRD
  patterns produced from equilibrated MD trajectories (see
  Section~\ref{method:xrd}) and comparing these simulated patters to experiment, addressing
  question~\ref{point:xrdmatch} in the introduction. We tested systems built with 5 columns per pore
  in the parallel displaced and sandwiched configurations at
  300 K in the ordered and disordered basins, as those were the most consistent
  with the most unambiguous XRD features, the pore-to-pore spacing
  and the stacking distance, as described in section~\ref{section:mon_per_pore}. 
  We generated simulated patterns using the locally
  equilibrated portion of each simulation trajectory (see
  Section~\ref{method:equil_time}). There are two important factors to take into
  account when looking at differences between experimental and simulated XRD patterns.
  First, the simulated diffraction patterns have some noise, especially along the
  $q_z$ axis at $q_r$=0, which is where several of the more interesting features
  are located. This is due to the angle averaging of the 3D structure factor
  around the $q_z$-axis, meaning there are fewer samples as $q_r$ approaches 0.
  More importantly, our simulations do not appear to be long enough to sample truly
  independent configurations within their respective metastable basins (see
  section~\ref{section:slow_dynamics}), meaning we must take care in interpreting
  the results of the XRD. The amount of independent configurations required to 
  converge along the $q_z$ axis is analyzed in Supporting Information, 
  Section S-\ref{S-section:xrd_noise}. The simulated patterns generated for all
  systems studied are shown and compared to experiment in Figure~\ref{fig:XRDsim}.

  \begin{figure}[!htb]
	\centering
	\begin{subfigure}{0.88\textwidth}
	\begin{subfigure}{0.28\linewidth}
			\begin{subfigure}{\textwidth}
			\centering
		        	\includegraphics[width=\linewidth]{rzplot_layered_300K_jet_nocbar.pdf}
			\end{subfigure}
			\begin{subfigure}{\textwidth}
		       		\centering
	        		\includegraphics[width=\linewidth]{rzplot_layered_300K_disorder_jet_nocbar.pdf}
			\end{subfigure}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
	\centering
			\begin{subfigure}{\textwidth}
		       		\centering
	        		\includegraphics[width=\linewidth]{WAXS_raw_jet_nocbar.pdf}
			\end{subfigure}
	\end{subfigure}
	\begin{subfigure}{0.28\linewidth}
	\centering
			\begin{subfigure}{\textwidth}
			\centering
		        	\includegraphics[width=\linewidth]{rzplot_offset_300K_jet_nocbar.pdf}
			\end{subfigure}

			\begin{subfigure}{\textwidth}
				\centering
			    \includegraphics[width=\linewidth]{rzplot_offset_300K_disorder_jet_nocbar.pdf}
			\end{subfigure}
	\end{subfigure}
	\end{subfigure}
	\begin{subfigure}{0.1\textwidth}
		\centering
		\includegraphics[width=\linewidth]{colorbar_jet.pdf}
	\end{subfigure}	
	\caption{Simulated XRD patterns show some qualitative agreement with
		experiment. Shown is a comparison of the (a) Sandwiched, ordered basin (b)
		Sandwiched, disordered basin (d) Parallel displaced, ordered basin and (e)
		Parallel displaced, disordered basin configurations with (c) experimental WAXS.
		The major reflections of interest (See Figure~\ref{fig:SWAXS} for their
		definitions) are present at varying degrees. In all cases, R-pores, R-alkanes
		and R-$\pi$ are present to some degree. R-spots is also present, however it is
		generally partially covered by the broad R-$\pi$ reflection. Quantitative
		comparisons of the relative intensities and locations of reflections of
		interest are presented in Table~\ref{table:relative_intensities_300K}. In both
		XRD patterns generated from parallel displaced patterns, there is a faint line
		across $q_z \sim 0.7~$\AA$^{-1}$, half the simulated value of R-$\pi$. Although
		it does not cross through $q_r = 0$~\AA$^{-1}$, it is at the same $q_z$ value
		where we expect R-double to be present. Still, R-double is not fully reproduced
		in any of the simulated patterns.}~\label{fig:XRDsim} 
  \end{figure}

@@

  Overall, the intensity of R-$\pi$ generated from the parallel displaced 
  ensemble shows the best agreement with experiment so far. For simplicity, 
  we will limit the following discussion to the parallel displaced configuration.
  Equivalent data for the sandwiched configuration is presented in Section 
  S\ref{S-section:sandwiched_ensemble}.

>>>>

  \begin{figure}[!htb]
        \centering
                \begin{subfigure}[t]{0.495\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated_words.pdf} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
	\begin{subfigure}[t]{0.38\linewidth}
                \centering
                \includegraphics[width=\textwidth]{SAXS.pdf}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
	\caption{(a) 2D-WAXS gives details about repeating features on the
		order of angstroms. Experimentalists have explained each of the 5 major
		reflections present as follows: (R-$\pi$) Aromatic head groups $\pi-\pi$ stack
		3.7 \AA~apart. (R-double) Monomers arrange vertically in a 2\textsubscript{1}
		helix. (R-alkanes) Alkane chain tails pack 4.5 \AA~apart. (R-spots) Monomer
		tails are tilted with respect to the membrane plane. (R-pores) As derived from
		SAXS, the pores are spaced 4.12 nm apart and pack hexagonally (b) (Reproduced
		from Ref.\citenum{feng_thin_2016}) The repeat spacing in the 1D-SAXS
		scattering pattern is characteristic of hexagonal packing. The leading peak,
		q*, represents the distance between the d\textsubscript{100} planes, which translates 
                into a distance between pore centers of 4.12 nm.} 
	\label{fig:SWAXS}
 \end{figure}

@@

  \subsection{Source code}

  Python scripts used to conduct post-simulation trajectory analysis, are available online at \\ 
  \texttt{https://github.com/bencoscia/llcsim}. The python scripts used to simulated XRD
  patterns are publicly available online at \texttt{https://github.com/joeyelk/MD-Structure-Factor}. 
  Additional details are available in Section \ref{S-section:python_scripts} of the SI.

@@

  We generated an ensemble of characteristic, low-energy vacuum monomer
  configurations by applying a simulated annealing process to a
  parameterized monomer. We cooled monomers from 1000 K to 50 K over 10
  ns. We randomly pulled a low energy configuration from the
  trajectory then reassigned charges using \texttt{molcharge}. Using the new
  charges, we annealed the monomer system again and pulled a random monomer
  configuration from the trajectory which we used for full system
  construction. Section \ref{S-section:parameterization} 
  of the Supporting Information (SI) provides further detail of the parameterization process.

@@

  The time scale for self-assembly of monomers into the hexagonal phase is
  unknown and likely outside of a reasonable length for an atomistic simulation,
  requiring a more efficient way to build the system. Previous work has shown
  a united-atom model of LLC can self-assemble into the H\textsubscript{II} phase
  configuration in $\sim$1000 ns \cite{mondal_self-assembly_2013}.  We
  attempted atomistic self-assembly by packing monomers into a box using Packmol
  \cite{martinez_packmol:_2009}. Simulations of greater than 100 ns show no
  indicators of progress towards an ordered system (see Section \ref{S-section:self_assembly} of the 
  SI). To bypass the slow self-assembly process, we used a 
  variety of automated procedures to assemble monomers into a structure close to
  one of a number of hypothesized equilibrium configurations (See Figure 
  \ref{S-fig:build_procedure} of the SI). These initial structures may result
  in equilibration into slowly-intercoverting metastable states, an issue we will
  address later in the paper.

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of columns of stacked monomers with periodic
  continuity along the pore axis, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. We prefer a small number of stacked
  monomers in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomers
  per column in order to minimize finite size effects with reasonable computational expense 
  and to obtain sufficient resolution when simulating XRD patterns (see further discussion
  in Section \ref{S-section:monomers_per_column} of the SI).

@@

  When constructing an initial configuration, there are a number of variables
  which require careful consideration while placing monomers. We find that the equilibrium
  configuration is sensitive to some while insensitive to others. For example, we find the starting
  pore radius, defined as the distance of a chosen head group carbon from the
  pore's central axis, does not influence the equilibrium structure if one chooses
  a reasonable value. The initial distance between pores, within a wide range, also has 
  little effect on the equilibrated structure. However, one should not start them
  too close or there will be high energy repulsions during early equilibration. We chose a
  pore radius of 0.5 nm and an initial pore spacing of 4.5 nm, $\sim$10\% larger than the 
  experimental value of 4.12 nm for our initial configurations. A sensitivity analysis of 
  both parameters is presented in the SI, Section \ref{S-section:initial_config_dependence}. The 
  distance between vertically stacked monomers, the xy position of monomers with respect 
  to vertically adjacent monomers, and the number of columns per pore do influence the 
  equilibrated structure and require further justification for their choices. We rely on 
  experimental data to inform them.

  We chose the vertical spacing between monomers for the initial configuration based
  on R-$\pi$ and then allowed the system to readjust during equilibration. We rotated 
  each monomer so the plane of its aromatic head group would be coplanar with the xy plane. We
  explored three different initial monomer spacings. The first is exactly
  equal to R-$\pi$ with monomers placed so aromatic rings stack 3.7 \AA~apart in
  the $z$-direction. We also extensively explored a second system with an initial spacing of 5
  \AA. We briefly explored a third system with an initial spacing of 10
  \AA. However this largest spacing yields non-physical behavior which is detailed in the 
  SI, Section \ref{S-section:initial_dbwl}.

  We chose the relative orientation between vertically adjacent monomers in each column 
  based on clues from diffraction data as well as the various known stacking modes of 
  benzene and substituted benzene rings: sandwiched, parallel-displaced and T-shaped
  \cite{sinnokrot_estimates_2002}. We ruled out the T-shaped configuration
  because its $\sim$5 \AA~equilibrium stacking distance \cite{sinnokrot_estimates_2002}
  is inconsistent with R-$\pi$. It is also infeasible for the monomers to orient in the 
  T-shaped conformation because of the bulky tail groups. We explored the system's 
  preference towards the sandwiched vs. parallel displaced stacking modes in some detail.
  Both have reported stacking distances near the R-$\pi$ value of 3.7 \AA. Head groups in
  our sandwiched initial configuration stack directly on top of each other while
  head groups in the parallel displaced initial configuration stack with an offset
  of $180^\circ/ncol$ where $ncol$ is the number of columns per pore. See Figure 
  \ref{S-fig:stacking} of the SI for a detailed illustration 
  of the initial configurations in each mode.

@@

  We developed equilibration schemes for creating dry (i.e. thermotropic LC phase)
  and wet (i.e. LLC phase) configurations.
  Both schemes start with an initial configuration generated according to the
  previous guidelines. For wet systems, we added the desired concentration of
  water to the initial configuration and carried out equilibration in the same
  way as the dry systems. First, we fixed monomer head groups in place using
  position restraints with a force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We
  gradually released the position restraints by decreasing the force constants
  over a series of NVT simulations. We allowed the resulting unrestrained
  structure to equilibrate for 5 ns in the NPT ensemble with pressure controlled
  by the Berendsen barostat, followed by NPT equilibration simulations run for at
  least 400 ns using the Parrinello-Rahman barostat. More equilibration details
  are given in Section \ref{S-section:equilibration} of the SI.

@@

  To calculate the equilibrated pore spacing, we measured the distance between
  pore centers. We located the pore centers by averaging the coordinates of sodium
  ions in their respective pores. We generated pore spacing statistics 
  using the bootstrapping technique (See Section \ref{S-section:p2p_stats} of the
  SI). 
  The pore spacings calculated in this way are consistent
  with one half of the $x$ and $y$ box vectors (See, Table \ref{S-table:p2p} in the
  SI). However, our method for their calculation does a better 
  job capturing the spread of pore-to-pore distances.

@@

  We measured the intensity of R-pores based on the intensity of the
  d\textsubscript{100} peak (the leading peak closest to $q_r$ = 0) of the $q_r$
  cross-section of the patterns at $q_z$=0 (Figure~\ref{fig:saxs_xsection}). The
  beamstop covers most of the small angle reflections in the experimental 2D-WAXS
  pattern. In order to compare the simulated intensity of R-pores to experiment,
  we used the $q_r$ cross-section of 2D-SAXS which was generated from the same
  sample (Figure \ref{S-fig:2DSAXS} of the SI). Since the d\textsubscript{200} peak is
  partially exposed in the experimental 2D-WAXS pattern, we normalized the
  experimental 2D-SAXS cross-section by matching the intensity of the
  d\textsubscript{200} peak between it and the experimental 2D-WAXS
  cross-section. It is possible that the peak is not fully captured in the 2D-WAXS
  pattern and that we have underestimated the intensity of R-pores in the
  experimental pattern.

@@

  We explored the pores' compositions by measuring the average number densities
  of various monomer components as a function of distance from the pore centers.
  We looked at the average number density of sodium ions, aromatic rings and 
  carbon atoms making up the monomer tails. We binned the radial distance of all
  atoms in each group from the pore centers, then normalized by the volume of the
  annulus defined by the bin edges and the z box vector (See Figure \ref{S-fig:rdf_diagram}
  of the SI).

@@

  We calculated ionic conductivity using the Nernst-Einstein relationship, which 
  relates the DC ionic conductivity, $\sigma$, to ion diffusivity, $D$, 
  concentration, $C$, ion charge, $q$, the Boltzmann constant, $k_b$, and 
  absolute temperature, $T$: 
  \begin{equation}
	\sigma = \dfrac{q^2CD}{k_b T} 
	\label{eqn:nernst_einstein}
  \end{equation}
  We measured sodium ion diffusion coefficients by calculating the slope
  of the linear region of the $z$-direction mean square displacement curve as
  indicated by the Einstein relation \cite{einstein_investigations_1956}. We
  visualized the MSD plot to determine where to begin and end a linear fit. We
  measured ion concentration with respect to the volume of the entire unit cell. 
  More details are provided in the SI, Section \ref{S-section:ionic_conductivity}.

@@

  We based our cross-linking algorithm on the known reaction mechanism 
  (Figure \ref{S-fig:xlink_mech} of the SI). The reaction 
  takes place at the terminal vinyl groups
  on each alkane tail. The procedure is carried out iteratively. Each iteration, the
  algorithm chooses carbons to cross-link based on the distance between eligible 
  carbon pairs. The algorithm then updates the topology with the new bonds and atom
  types, energy minimizes the system and runs a short simulation before selecting 
  the next group of eligible carbons atoms.

@@

  Our simulations best support a model built with 5 monomer columns per pore based on
  the measured equilibrated pore-to-pore distances. To identify the density of
  monomer columns around each pore, addressing question \ref{point:monomernum} in the introduction, we ran
  simulations of systems created with 4--8 columns per pore. We built systems in both
  the parallel displaced and sandwiched configurations and equilibrated them according
  to the dry equilibration procedure. We tested all systems with an initial vertical monomer 
  spacing, $\mathit{d}$, of 3.7 \AA~in accordance with R-$\pi$. We tested 4 additional 
  systems with monomers initially spaced 5 \AA~apart vertically (see Section \ref{S-section:initial_pore_spacing}
  of the SI for more details on sensitivity to initial monomer spacing). 
  We considered the pore-to-pore spacing to be equilibrated as defined in Section~\ref{method:equil_time}.
  Figure~\ref{fig:p2p} shows the equilibrated pore-to-pore distances for all systems tested.

@@

  All systems tested, although equilibrated from the perspective of the metrics used here, are
  frozen in metastable basins. Not all make physical sense or fit the experimental profile that 
  we are trying to match. In the limit of infinite simulation time, all systems will in theory 
  converge to a single equilibrium configuration, but that time is far beyond the 100's of 
  nanosecond simulated here. For simplicity, we group the systems studied here into the ordered
  and disordered basins. What we find is that any system equilibrated with $\mathit{d}=3.7$ \AA~can generally 
  be characterized as being in a more ordered basin, and any system equilibrated with $\mathit{d}=5.0$ \AA~can
  be characterized as begin in a more disordered basin. The extra space between stacked monomers
  in the disordered basin systems gives the monomer head groups more rotational freedom. We 
  quantified the ordering of the head groups using the nematic order parameter (see Section 
  \ref{S-method:nematic_order} for details of the calculation). Disordered basin systems have
  a lower nematic order parameter (meaning they are more disordered) than ordered basin systems. 
  Generally, when monomers are started further apart, they stay further apart than systems where
  monomers are started closer together (see Table~\ref{table:correlation_length}). This increased
  equilibrated stacking distance of monomers in disordered basin systems increases their discrepancy
  with experimental stacking distances derived from WAXS relative to ordered basin systems 
  (See Section~\ref{section:rpi}).

@@

  We can better understand the structure of the system by simulating XRD
  patterns produced from equilibrated MD trajectories (see
  Section~\ref{method:xrd}) and comparing these simulated patters to experiment, addressing
  question~\ref{point:xrdmatch} in the introduction. We tested systems built with 5 columns per pore
  in the parallel displaced and sandwiched configurations at
  300 K in the ordered and disordered basins, as those were the most consistent
  with the most unambiguous XRD features, the pore-to-pore spacing
  and the stacking distance, as described in Section~\ref{section:mon_per_pore}. 
  We generated simulated patterns using the locally
  equilibrated portion of each simulation trajectory (see
  Section~\ref{method:equil_time}). There are two important factors to take into
  account when looking at differences between experimental and simulated XRD patterns.
  First, the simulated diffraction patterns have some noise, especially along the
  $q_z$ axis at $q_r$=0, which is where several of the more interesting features
  are located. This is due to the angle averaging of the 3D structure factor
  around the $q_z$-axis, meaning there are fewer samples as $q_r$ approaches 0.
  More importantly, our simulations do not appear to be long enough to sample truly
  independent configurations within their respective metastable basins (see
  Section~\ref{section:slow_dynamics}), meaning we must take care in interpreting
  the results of the XRD. The amount of independent configurations required to 
  converge along the $q_z$ axis is analyzed in SI, 
  Section \ref{S-section:xrd_noise}. The simulated patterns generated for all
  systems studied are shown and compared to experiment in Figure~\ref{fig:XRDsim}.

  \begin{figure}[!htb]
	\centering
	\begin{subfigure}{0.905\textwidth}
	\begin{subfigure}{0.28\linewidth}
			\begin{subfigure}{\textwidth}
			\centering
		        	\includegraphics[width=\linewidth]{rzplot_layered_300K_jet_nocbar.pdf}
			\end{subfigure}
			\begin{subfigure}{\textwidth}
		       		\centering
	        		\includegraphics[width=\linewidth]{rzplot_layered_300K_disorder_jet_nocbar.pdf}
			\end{subfigure}
	\end{subfigure}
	\begin{subfigure}{0.4\linewidth}
	\centering
			\begin{subfigure}{\textwidth}
		       		\centering
	        		\includegraphics[width=\linewidth]{WAXS_raw_jet_nocbar.pdf}
			\end{subfigure}
	\end{subfigure}
	\begin{subfigure}{0.28\linewidth}
	\centering
			\begin{subfigure}{\textwidth}
			\centering
		        	\includegraphics[width=\linewidth]{rzplot_offset_300K_jet_nocbar.pdf}
			\end{subfigure}

			\begin{subfigure}{\textwidth}
				\centering
			    \includegraphics[width=\linewidth]{rzplot_offset_300K_disorder_jet_nocbar.pdf}
			\end{subfigure}
	\end{subfigure}
	\end{subfigure}
	\begin{subfigure}{0.085\textwidth}
		\centering
		\includegraphics[width=\linewidth]{colorbar_jet.pdf}
	\end{subfigure}	
	\caption{Simulated XRD patterns show some qualitative agreement with
		experiment. Shown is a comparison of the (a) Sandwiched, ordered basin (b)
		Sandwiched, disordered basin (d) Parallel displaced, ordered basin and (e)
		Parallel displaced, disordered basin configurations with (c) experimental WAXS.
		The major reflections of interest (See Figure~\ref{fig:SWAXS} for their
		definitions) are present at varying degrees. In all cases, R-pores, R-alkanes
		and R-$\pi$ are present to some degree. R-spots is also present, however it is
		generally partially covered by the broad R-$\pi$ reflection. Quantitative
		comparisons of the relative intensities and locations of reflections of
		interest are presented in Table~\ref{table:relative_intensities_300K}. In both
		XRD patterns generated from parallel displaced patterns, there is a faint line
		across $q_z \sim 0.7~$\AA$^{-1}$, half the simulated value of R-$\pi$. Although
		it does not cross through $q_r = 0$~\AA$^{-1}$, it is at the same $q_z$ value
		where we expect R-double to be present. Still, R-double is not fully reproduced
		in any of the simulated patterns.}~\label{fig:XRDsim} 
  \end{figure}

@@

  Overall, the intensity of R-$\pi$ generated from the parallel displaced 
  ensemble shows the best agreement with experiment so far. For simplicity, 
  we will limit the following discussion to the parallel displaced configuration.
  Equivalent data for the sandwiched configuration is presented in Section 
  \ref{S-section:sandwiched_ensemble}.
